---
title: "R Notebook"
output: html_notebook
---



```{r}
```{r}
#Upload packages
library(tidyverse)
library(readxl)
library(here)
library(factoextra)
library(ggdendro)
library(dendextend)
library(colormap)
library(indicspecies)
library(fuzzySim)
library(cluster)
library(patchwork)
library(vegan)
library(RColorBrewer)

#Need to go through and test these

```

```{r}
#Upload data

#Upload data from my master data standardization sheet
micro <- read_csv(here("outputs", "micro_master_2021-10-18.csv")) 

clust_phy <- read_csv(here("outputs", "clusters_phys.csv"))

clust_chem <- read_csv(here("outputs", "chemtax_clusters.csv"))
```

```{r}
micro <- micro %>% 
  filter(!site_id == "QU39")
```

```{r}
#Selecting the species groups I will work with and limiting species that rarely observed.

#For now, I am keeping all of the heterotrophic species (just in this workbook)

# micro <- micro %>% 
#   filter(trophicStatus == "auto")

#Counting how many times each species is observed
micro <- micro %>%
  group_by(scientificName) %>%
  mutate(num_occurrence = n()) %>% 
  ungroup()

#removing species that have not been observed at least twice -  this is to minimize the influence of zeros. When I used this, it removes some species that are important, but not observed often. So trying with it turned off.
# micro <- micro %>%
#   filter(num_occurrence > 5)

#Removing cyanobacteria as their counts are not reliable
micro <- micro %>% 
  filter(!scientificName_accepted == "Cyanobacteria")

micro <- micro %>% 
  filter(!group == "Metazoa") %>% 
  filter(!group == "Protozoa")
```

```{r}
#Trying to look at only species present in 20% of samples
#How many samples
sample_num <- micro %>% 
  distinct(date, site_id)

species_10 <- micro %>% 
  group_by(scientificName) %>% 
  summarise(n_obs = n()) %>%
  ungroup() %>% 
  mutate(perc_obs = n_obs/50) %>% 
  filter(perc_obs >= 0.10)

species_10_list <- species_10$scientificName

micro <- micro %>% 
  filter(scientificName %in% species_10_list)

```

```{r}
#Sub-setting the species determined via indicator species analysis on station
micro_ind_stat <- micro %>% 
  filter(scientificName %in% ind_list)

#Determining number of times each indicator species observed at each station
micro_ind_stat <- micro_ind_stat %>% 
  group_by(site_id, scientificName) %>% 
  summarise(n_obs = n(),
            max_count = max(species_sum),
            mean_count = mean(species_sum),
            median_count = median(species_sum)) %>% 
  ungroup() 

test <- micro %>% 
  filter(!site_id == "QU39" & class  == "Bacillariophyceae") %>% 
  group_by(site_id, scientificName) %>% 
  summarise(n_obs = n(),
            max_count = max(species_sum),
            mean_count = mean(species_sum),
            median_count = median(species_sum)) %>%
  ungroup()

#Maybe Friday brain, but this is the best way I could find to count how many times each station was sampled - Want this so I can determine percentage of times each indicator species was observed by station
n_station <- micro %>% 
  filter(!site_id == "QU39") %>% 
  group_by(date, site_id) %>% 
  distinct(date, site_id) %>% 
  ungroup() %>% 
  group_by(site_id) %>% 
  summarise(n_samp = n()) %>% 
  ungroup()

#Joining total number of times each station was sampled with stats on how many times each species was observed at each station.
micro_ind_stat <- micro_ind_stat %>% 
  left_join(n_station) %>% 
  mutate(perc_obs = n_obs/n_samp,
         label = scales::percent(perc_obs %>% round(2)))
```

```{r}
#Creating a placeholder for Ceratium lineatum for DFO2 - makes plot save a space so the groupings stay consistent.
dfo_cl <- data.frame(site_id = "DFO2",
                     scientificName = "Ceratium lineatum",
                     n_obs = 0,
                     max_count = 0,
                     mean_count = 0,
                     median_count = 0,
                     n_samp = 0,
                     perc_obs = 0,
                     label = "0%")
                 
micro_ind_stat <- rbind(dfo_cl, micro_ind_stat)
```

```{r}
#Creating a plot showing number and percent observed for each indicator species by station. Then a second plot with the average abundance of each species for each station.

p1 <- micro_ind_stat %>% 
  filter(!scientificName == "Skeletonema marinoi" |
         !scientificName == "Thalassiosira rotula" |
         !scientificName == "Chaetoceros seiracanthus") %>% 
  ggplot(aes(x = perc_obs, y = scientificName, fill = site_id)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), 
           color = "black") +
  scale_fill_brewer(palette = "Set1") +
  scale_x_continuous(expand = c(0.01, 0), limits = c(0, 1.03)) +
  scale_y_discrete(labels = c("Chaetoceros debilis" = "C.deb.",
                              "Chaetoceros decipiens" = "C.dec.",
                              "Chaetoceros seiracanthus" = "C.s.",
                              "Dactyliosolen fragilissimus" = "D.f.",
                              "Dactyliosolen phuketensis" = "D.ph.",
                              "Detonula pumila" = "D.p.",
                              "Guinardia delicatula" = "G.d.",
                              "Rhizosolenia setigera" = "R.s.",
                              "Skeletonema marinoi" = "S.m.",
                              "Thalassiosira rotula" = "T.r.",
                              "Ceratium lineatum" = "C.l.")) +
  theme_bw() +
  labs(x = "% Observed") +
  theme(text = element_text(size = 25),
        axis.title.y = element_blank(),
        axis.text = element_text(colour = "black"),
        legend.position = "none")
  

p2 <- micro_ind_stat %>% 
  filter(!scientificName == "Skeletonema marinoi" |
         !scientificName == "Thalassiosira rotula" |
         !scientificName == "Chaetoceros seiracanthus") %>% 
  ggplot(aes(x = mean_count, y = scientificName, fill = site_id)) +
  geom_bar(stat = "identity", position = position_dodge2(preserve = "single"), 
           color = "black") +
  scale_fill_brewer(palette = "Set1") +
  # scale_x_continuous(expand = c(0, 1000)) +
  # ggbreak::scale_x_break(c(55000, 240000)) +
  labs(x = "Mean Abundance (cells l)") +
  scale_x_continuous(expand = c(0.01, 0), limits = c(0, 26000)) +
  scale_y_discrete(labels = c("Chaetoceros debilis" = "C.deb.",
                              "Chaetoceros decipiens" = "C.dec.",
                              "Chaetoceros seiracanthus" = "C.s.",
                              "Dactyliosolen fragilissimus" = "D.f.",
                              "Dactyliosolen phuketensis" = "D.ph.",
                              "Detonula pumila" = "D.p.",
                              "Guinardia delicatula" = "G.d.",
                              "Rhizosolenia setigera" = "R.s.",
                              "Skeletonema marinoi" = "S.m.",
                              # "Thalassiosira rotula" = "T.r.",
                              "Ceratium lineatum" = "C.l.")) +
  theme_bw() +
    theme(text = element_text(size = 25),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.text = element_text(colour = "black"),
          legend.position = c(0.85, 0.1),
          legend.title = element_blank())

fig <- p1 + p2

ggsave(here("figures_good", "indicator_species_station_bar.png"), fig,
       width = 12, height = 7, dpi = 300)
```


